---
title: "Forecast daily electricity demand"
author: "Biz Yoder & Ryan McCord"
date: "2023-04-02"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

```{r setup, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE, fig.align='center', echo=FALSE, include=FALSE}
#Import libraries
library(readxl) #import excel
library(dplyr) #pipes
library(forecast) #time series
library(ggplot2)
```

```{r, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE, fig.align='center'}
#Import load data
load <- read_excel(path="./Data/load.xlsx") 
```

```{r, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE, fig.align='center'}
#Find daily load (average of hourly values)
load_daily <- load %>%
  mutate(daily_load = rowMeans(select(load, starts_with('h')), na.rm = TRUE))
```

```{r, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE, fig.align='center'}
#Construct time series object
ts_daily <- msts(load_daily$daily_load, start = c(2005, 01, 01), end = c(2010, 12, 31),seasonal.periods = c(7,365.25))
```

## Modeling and Forecasting with complex seasonalities

We learned some approaches to model series with a single seasonal pattern.

* Moving Average, Naive Methods
* Simple Exponential smoothing
* Seasonal ARIMA
* State Space models
* ETS - Exponential smoothing state space model

Today we will learn a few examples of models for multiple seasonal patterns.

* ETS - another version of exponential smoothing (with 2 seasonal components)
* ARIMA with dynamic harmonic fourier components - ARIMA model that can deal with multiple seasonal components
* TBATS 
* Neural Network
```{r, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE, fig.align='center'}
#Visualize data
summary(load_daily$daily_load)

ggplot(load_daily, aes(x=load_daily$date,y=load_daily$daily_load)) +
  geom_line() +
  xlab("Date") +
  ylab("Average Daily Household Power Demanded")
  
ts_daily %>% mstl() %>%
  autoplot()
```


```{r, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE, fig.align='center'}
#Separate into train and testing data

load_daily_train <- load_daily$daily_load[1:1826]
train_dates <- as.Date(load_daily$date[1:1826])
load_daily_train <- cbind.data.frame(train_dates, load_daily_train)
load_daily_test <- load_daily$daily_load[1827:2191]
test_dates <- as.Date(load_daily$date[1827:2191])
load_daily_test <- cbind.data.frame(test_dates, load_daily_test)

#what is going on here???
ts_daily_train <- msts(load_daily_train, start = c(2005, 01, 01), end = c(2009, 12, 31),seasonal.periods = c(7,365.25))

ts_daily_test <- msts(load_daily_test, start = c(2010, 01, 01), end = c(2010, 12, 31),seasonal.periods = c(7,365.25)) #something weird with seasonality #s


autoplot(ts_daily_train)
autoplot(ts_daily_test)
autoplot(ts_daily)
```

